---
title: "Peer Assessment 1"
author: "Celso Castro"
date: "Thursday, July 16, 2015"
output: html_document
---

First we read the data

```{r readdata}
datos=read.csv("C:/Users/Celso/Documents/activity.csv",sep=",",header=T)
```
Next, we calculate the total number of steps taken per day

```{r countdailysteps}
totalpasos<- tapply(datos$steps, datos$date, sum, na.rm = TRUE)
```

This histogram shows the frequency of daily steps taken.

```{r totalNA}
complete<-complete.cases(datos)
datos2=datos[complete,]
tNA<-nrow(datos)-nrow(datos2)
```

The total number of (missing values) NA's in the database is `r tNA`

```{r histogramdailysteps}
hist(totalpasos,main="Número de pasos por día", xlab="Pasos por día",ylab="Frecuencia",breaks=30,border="blue")
```

```{r meanmedian}
prom<-round(mean(totalpasos),0)

mediana<-median(totalpasos)
```

The mean of the total steps taken is `r prom` and the median is `r mediana`

```{r stepsbyinterval}
pasosint<- tapply(datos$steps, datos$interval, mean, na.rm = TRUE)
plot(levels(as.factor(datos$interval)),pasosint,main="Promedio de pasos por intervalo", xlab="Intervalo de 5 minutos",ylab="Promedio de pasos hechos",type="l",col="blue")
```


```{r maximumstepsbyinterval}
maximo=max(pasosint)
inter<-names(pasosint)[which(pasosint == max(pasosint))]
```

The interval which contains the maximum number of steps is `r inter`

```{r FillNAswithMean}

        datosnoNA<-datos
        for (i in 1:nrow(datos)) {

        if (is.na(datos$steps[i])==TRUE) {

            intervalo <- as.character(datos$interval[i])
            mediaint <-round(pasosint[intervalo],0)

            datosnoNA$steps[i] <- mediaint
                                         }
                                  }         
head(datosnoNA)
```

The data set datosnoNA has no missing values

```{r histogramdailystepsfilleddata}
totalpasos2<- tapply(datosnoNA$steps, datosnoNA$date, sum,na.rm=TRUE)
hist(totalpasos2,main="Número de pasos por día", xlab="Pasos por día",ylab="Frecuencia",breaks=30,border="blue",ylim=c(0,20))
prom2<-round(mean(totalpasos2),0)
mediana2<-median(totalpasos2)
```

The mean of the total steps taken without missing values is `r sprintf("%0.0f",prom2)` and the median is `r  sprintf("%0.0f",mediana2)`

*We can conclude that when the missing observations are changed for the meand of steps by interval there is a significant increase in the dayly steps taken which have a direct impact on the mean and median of these variable. These values are bigger than the original ones.

```{r factorweekdayweekend}
datosnoNA$date<-as.Date(datosnoNA$date)
dias<-as.factor(weekdays(datosnoNA$date))
datosnoNA$dias<-dias
#we have to check that is a weekday or a weekend and rename the factors in the order they are
levels(datosnoNA$dias) #to check the order
levels(datosnoNA$dias)<-c("weekend","weekday","weekday","weekday","weekday","weekend","weekday")
```

Now we generate the plane plot
```{r plotweekdayweekend}
library(plyr)
library(ggplot2)
qplot(datosnoNA$interval,datosnoNA$steps, data=arrange(datosnoNA, date), facets= dias~., geom="line", xlab="Intervalo", ylab="Número de pasos",main="Promedio de pasos por días de semana y fin de semana")
```

